---
title: "All3Trial Olabisi"
author: "Olabisi Adegbesan"
date: '2022-07-11'
output: 
  word_document:
    reference_docx: RtemplateCQ.docx
    toc: yes
---

```{r setup, include=FALSE}
# base size Plot
#plotsize <- 3
#knitr::opts_chunk$set(
  #echo = FALSE,
  #fig.height = 0.7 * plotsize,
  #fig.width = plotsize,
 # message = FALSE,
  #warning = FALSE,
  #comment = NA,
  #dpi = 300,
  #results = "asis")
pacman::p_load(conflicted,wrappedtools,nlme,broom,
               multcomp,tidyverse,foreign,DescTools, ez,
               ggbeeswarm, readxl, plotrix, ggsignif, rlist, here, skimr, flextable, ftExtra, Hmisc, table1, nlme, patchwork, aomisc, deSolve, flextable, officer)#conflicted,
# rayshader,av)
pacman::p_unload(DescTools, foreign)
# conflict_scout()
conflict_prefer('select','dplyr')
conflict_prefer('filter','dplyr')

#set chunks defaults
knitr::opts_chunk$set(
  echo       = FALSE,
  message    = FALSE,
  warning    = FALSE,
  comment = NA,
  fig.width = 7,
  fig.asp = 0.8,
  out.width = "80%",
  dpi = 300
)

# set flextable defaults
knitr::opts_chunk$set(echo = FALSE, fig.cap = TRUE)
  set_flextable_defaults(
  font.family = "Arial", font.size = 9, 
  theme_fun = "theme_vanilla",
  big.mark="", table.layout="autofit")
  
# formatting properties for specific paragraphs
centeredP <- fp_par(text.align = "center")


```

# Information for the datasets 
## Dosing
For all three trials, oral bolus doses were administered to subjects and dose amount administered can be found in the column DoseAMT (unit: mg)

## Observations
**StudyID**: Trial ID
**ID**: subject ID
**Time (hour)**: the observation time points. 
**C1Obs (mg/L)**: observation for plasma parent drug concentration
**CmObs (mg/L)**: observation for plasma metaolite concentration
**AuObs (mg)**: observation for parent drug in urine 
**PDRObs**: observation for PD effect 
**Note**: BQL value for parent drug is chosen to be 6.5e-3, and the one for metabolite is 6.5e-4.

## Covariates
**Sex**: 0 denotes male, and 1 denotes female
**Age** (years)
**Height (m)**
**Weight (kg)**
**Scr (mg/100ml)**: serum creatinine 
**ESRD**: indicator of whether the subject is in ESRD stage or not (1 means yes, 0 means no)
**Alb (g/100ml)**: serum albumin
**LabID**: the ID for the lab to which a subject is assigned to 
**nDrugs**: number of drugs taken by the subject
**nDiseases**: number of diseases that a subject has

# Import and Cleaning Data 
```{r import_strucrure}

all3trial <- read.csv("Data/All_3_study.csv") |>
  mutate(C1Obs=str_replace(C1Obs,'BQL', '6.5e-3'),
    CmObs=str_replace(CmObs,'BQL', '6.5e-4'),
    across(.cols = c(C1Obs:PDRObs,Height:Scr,Alb),
                .fns = as.numeric) ,
    Sex=factor(Sex,
               levels=c(0,1),
               labels=c('male','female')))

#all3trial
#glimpse(all3trial)

timed_data <- FindVars(varnames = 'Obs',allnames = cn(all3trial))
state_data <- FindVars(c('eight','Scr','Alb'), allnames = cn(all3trial))
```
## Visualization of all variables
```{r plot_of_all_variables}
all3trial |>
  filter(!is.na(Sex)) |> 
  pivot_longer(cols = c(C1Obs:PDRObs,Height:Scr,Alb)) |>
  ggplot(aes(value, fill=factor(Sex)))+
  geom_density(alpha=.2)+
  facet_wrap(facets = vars(name),
             scales = 'free')

cat('&nbsp;\n\n')
```

# Test for Normal distribution  
```{r kstests}
safe_ks <- purrr::possibly(ksnormal,otherwise = NA_real_)
safe_shapiro <- possibly(shapiro.test,
                         otherwise = list('p.value'=NA_real_))
```
## Static variables
```{r static_variables}
all3trial |>
  filter(Time==0) |> #since the state_data cannot change for the other time points
  group_by(Sex, StudyID) |> 
  summarise(
    across(.cols = state_data$names,
           .fns = ~safe_ks(.x) |> formatP())) |> 
  flextable() |> 
  set_caption('Results from KS-Test static variables') |> 
  set_table_properties(width=1,layout='autofit') |> 
  theme_zebra()
cat('&nbsp;\n\n')
cat('&nbsp;\n\n')

```
## Timed variables
```{r Timed_variables}
##StudyID 1
all3trial |> 
  filter(StudyID==1) |>
  group_by(Sex,Time) |> 
  summarise(
    across(.cols = timed_data$names,
           .fns = ~safe_shapiro(.x)$p.value |> formatP())) |> 
  arrange(Time) |> 
  flextable() |> 
  set_caption('Results from Shapiro-Test timed variables in Study1') |> 
  set_table_properties(width=1,layout='autofit') |> 
  theme_zebra()
cat('&nbsp;\n\n')

##StudyID 2
all3trial |> 
  filter(StudyID==2) |> 
  group_by(Sex,Time) |> 
  summarise(
    across(.cols = timed_data$names,
           .fns = ~safe_shapiro(.x)$p.value |> formatP())) |> 
  arrange(Time) |> 
  flextable() |> 
  set_caption('Results from Shapiro-Test timed variables in Study2') |> 
  set_table_properties(width=1,layout='autofit') |> 
  theme_zebra()
cat('&nbsp;\n\n')

##StudyID 3
all3trial |> 
  filter(StudyID==3) |> 
  group_by(Sex,Time) |> 
  summarise(
    across(.cols = timed_data$names,
           .fns = ~safe_shapiro(.x)$p.value |> formatP())) |> 
  arrange(Time) |> 
  flextable() |> 
  set_caption('Results from Shapiro-Test timed variables in Study3') |> 
  set_table_properties(width=1,layout='autofit') |> 
  theme_zebra()
cat('&nbsp;\n\n')
cat('\\newpage\n\n')
gaussvars <- FindVars(c('eight','Scr','Alb'), allnames = cn(all3trial))
ordvars <- FindVars('Obs', allnames = cn(all3trial))
  qualvars <- FindVars(c('ESRD','nD', 'Sex'), allnames = cn(all3trial))
```
# Group comparisons  
## Normal  
```{r gaussvars_analysis}
#static variables
all3trial |> 
  filter(Time==0) |> 
  group_by(Sex) |>
  summarise(across(.cols = gaussvars$names,
                   .fns = list(n=~length(na.omit(.x)),
                               MeanSD=meansd,
                               SEM=~SEM(.x) |> roundR()),
                   .names = "{.col}__{.fn}"),
            .groups = 'drop') %>% 
  mutate_if(.predicate = is.numeric,.funs = as.character) %>% 
  pivot_longer(cols = -Sex, 
               names_to=c('Variable','what'),
               values_to = 'Statistic',
               names_pattern='(.+)__(.+)') %>% 
  pivot_wider(names_from = c(what,Sex),
              names_glue = '{what} ({Sex})',
              values_from=Statistic) |>
  flextable() |> 
  set_caption('Group description for sex in variables with assumed Gaussian distribution') |> 
  set_table_properties(width=1,layout='autofit') |> 
  theme_zebra()
cat('&nbsp;\n\n')
cat('&nbsp;\n\n')
all3trial |> 
  filter(Time==0) |> 
  compare2numvars(
    dep_vars = gaussvars$names,
    indep_var = 'Sex',
    gaussian = TRUE) |> 
  # select(-p) |> 
  flextable() |> 
  set_caption('Group comparison for sex in variables with assumed Gaussian distribution') |> 
  set_table_properties(width=1,layout='autofit') |> 
  theme_zebra()
cat('&nbsp;\n\n')
cat('&nbsp;\n\n')
```

## Ordinal  
```{r ordvars_analysis}
safe_mclb <- possibly(median_cl_boot,
                      otherwise = tibble(median=NA_real_,
                                         CIlow=NA_real_,
                                         CIhigh=NA_real_))
all3trial |> 
  filter(Time==1) |> #chose time=1hr for the descriptive statistics
  group_by(Sex) |>
  summarise(across(.cols = ordvars$names,
                   .fns = list(n=~length(na.omit(.x)),
                               Desc=median_quart,
                               CI_Median=~safe_mclb(.x) |>
                                 select(-1) |> 
                                 roundR() |> 
                                 paste(collapse = ' .. ')),
                   .names = "{.col}__{.fn}"),
            .groups = 'drop') %>% 
  mutate_if(.predicate = is.numeric,.funs = as.character) %>% 
  pivot_longer(cols = -Sex, 
               names_to=c('Variable','what'),
               values_to = 'Statistic',
               names_pattern='(.+)__(.+)') %>% 
  pivot_wider(names_from = c(what,Sex),
              names_glue = '{what} ({Sex})',
              values_from=Statistic) |>
  mutate(across(everything(),
                ~str_replace(.x,
                             pattern = 'NA .. NA',
                             replacement = 'no data'))) |> 
  flextable() |> 
  set_caption('Group description for sex in ordinal variables') |> 
  set_table_properties(width=1,layout='autofit') |> 
  theme_zebra()
cat('&nbsp;\n\n')
cat('&nbsp;\n\n')

all3trial |> 
  filter(Time==1) |> 
  compare2numvars(
    dep_vars = ordvars$names,
    indep_var = 'Sex',
    gaussian = FALSE)|> 
  flextable() |> 
  set_caption('Group comparison for sex in ordinal variables') |> 
  set_table_properties(width=1,layout='autofit') |> 
  theme_zebra()
cat('&nbsp;\n\n')
cat('&nbsp;\n\n')
```

## Categorical  
```{r qualvars_analysis}
all3trial |> 
  filter(Time==0) |> 
  compare2qualvars(
    dep_vars = qualvars$names[-1],
    indep_var = 'Sex',
    spacer = ' ') |> 
  flextable() |> 
  set_caption('Group comparison for sex in factorial variables') |> 
  set_table_properties(width=1,layout='autofit') |> 
  flextable::align(~p==' ',j = 1,
                   align = 'center') |>    
  flextable::bg(~p!=' ',bg = 'lightgrey') 
cat('&nbsp;\n\n')
cat('\\newpage\n\n')

```
# Plasma Parent Drug Concentration-Time profile per Study
```{r C1Obs}
#all3trial_obs <- all3trial |> 
  #mutate(StudyID=factor(StudyID,
               #levels=c(1,2,3),
               #labels=c('Study1','Study2', 'Study3')))

#all3trial_obs |>
#ggplot(aes(x = Time, y = C1Obs, color = StudyID)) +
  #geom_boxplot()

 #grouping age into 4 bins
all3trial$AgeBins <- cut(all3trial$Age, 4)
#table(all3trial$AgeBins)

cat('&nbsp;\n\n')


```
## C1Obs-Time Profile for Study 1
```{r Study1}
#Study1
p1 <- all3trial |> 
  filter(StudyID==1, ID %in% c(1:7)) |> 
  #ggplot(aes(x=Time, y=C1Obs , colour = StudyID)) +
  ggplot(aes(x=Time, y=C1Obs , colour = factor(ID))) +
  geom_point(size= 3, alpha=0.3)+
  geom_line(size = .5)+
  #facet_wrap(~ID)+
  theme_minimal()+
  xlab("Time (Hours)") + ylab("Plasma Concentration (mg/L)") +
  ggtitle("Plasma Parent Drug Concentration - \n Time Profile (21 mg)")

p2 <- all3trial |> 
  filter(StudyID==1, ID %in% c(8:14)) |> 
  #ggplot(aes(x=Time, y=C1Obs , colour = StudyID)) +
  ggplot(aes(x=Time, y=C1Obs , colour = factor(ID))) +
  geom_point(size= 3, alpha=0.3)+
  geom_line(size = .5)+
  #facet_wrap(~ID)+
  theme_minimal()+
  xlab("Time (Hours)") + ylab("Plasma Concentration (mg/L)") +
  ggtitle("Plasma Parent Drug Concentration - \n Time Profile (42 mg)")

p3 <- all3trial |> 
  filter(StudyID==1, ID %in% c(14:21)) |> 
  #ggplot(aes(x=Time, y=C1Obs , colour = StudyID)) +
  ggplot(aes(x=Time, y=C1Obs , colour = factor(ID))) +
  geom_point(size= 3, alpha=0.3)+
  geom_line(size = .5)+
  #facet_wrap(~ID)+
  theme_minimal()+
  xlab("Time (Hours)") + ylab("Plasma Concentration (mg/L)") +
  ggtitle("Plasma Parent Drug Concentration -Time Profile (84 mg)")

(p1+p2)/p3

#with these plots, we can extrapolate the Cmax, Tmax, AUC and Therapeutic range.

cat('&nbsp;\n\n')
cat('&nbsp;\n\n')

```
## C1Obs-Time Profile for Study 2
```{r Study2}
#Study 2
##Individuals without ESRD (=0)
r1 <- all3trial |> 
  filter(StudyID==2, ESRD ==0) |> 
  ggplot(aes(x=Time, y=C1Obs , colour = factor(ID))) +
  geom_point(size= 3, alpha=0.3)+
  geom_line(size = .5)+
  theme_minimal()+
  xlab("Time (Hours)") + ylab("Plasma Concentration (mg/L)") +
  ggtitle("Individuals without ESRD C1Obs-Time Profile (70 mg)")

# Individuals with ESRD (=1)
r2 <- all3trial |> 
  filter(StudyID==2, ESRD ==1) |> 
  ggplot(aes(x=Time, y=C1Obs , colour = factor(ID))) +
  geom_point(size= 3, alpha=0.3)+
  geom_line(size = .5)+
  theme_minimal()+
  xlab("Time (Hours)") + ylab("Plasma Concentration (mg/L)") +
  ggtitle("Individuals with ESRD C1Obs-Time Profile (70 mg)")

r1/r2
cat('&nbsp;\n\n')
cat('&nbsp;\n\n')

```
## C1Obs-Time Profile for Study 3
```{r Study3}
#Study 3
t1 <- all3trial |> 
  filter(StudyID==3, DoseAMT==0, Time %in% c(0, 24, 48, 72, 96, 120, 144)) |> 
  ggplot(aes(x=Time, y=C1Obs,colour = factor(ID))) +
  geom_point(size= 3, alpha=0.3)+
  geom_line(size = .5)+
  theme_minimal()+
  xlab("Time (Hours)") + ylab("Plasma Concentration (mg/L)") +
  ggtitle("Placebo Concentration -Time Profile")
cat('&nbsp;\n\n')
t2 <- all3trial |> 
  filter(StudyID==3, DoseAMT==130, Time %in% c(0, 24, 48, 72, 96, 120, 144)) |> 
  ggplot(aes(x=Time, y=C1Obs, colour = factor(ID))) +
  geom_point(size= 3, alpha=0.3)+
  geom_line(size = .5)+
  xlab("Time (Hours)") + ylab("Plasma Concentration (mg/L)") +
  ggtitle("Plasma Parent Drug Concentration - \n Time Profile (130 mg)")

  #t2 + facet_wrap(facets = vars(Sex))
  t2 + facet_wrap(~Sex, ncol=2) +
  ggtitle("Plasma Parent Drug Concentration - \n Time Profile by Sex (130 mg)")
  t2 + facet_wrap(~factor(AgeBins), ncol=2)  +
  ggtitle("Plasma Parent Drug Concentration - \n Time Profile by Age (130 mg)")
cat('&nbsp;\n\n') 
t3 <- all3trial |> 
  filter(StudyID==3, DoseAMT==260, Time %in% c(0, 24, 48, 72, 96, 120, 144)) |> 
  ggplot(aes(x=Time, y=C1Obs, colour = factor(ID))) +
  geom_point(size= 3, alpha=0.3)+
  geom_line(size = .5)+
  xlab("Time (Hours)") + ylab("Plasma Concentration (mg/L)") +
  ggtitle("Plasma Parent Drug Concentration - \n Time Profile (260 mg)")

  t3 + facet_wrap(~Sex, ncol=2) +
  ggtitle("Plasma Parent Drug Concentration - \n Time Profile by Sex (260 mg)") 
  t3 + facet_wrap(~factor(AgeBins), ncol=2) +
  ggtitle("Plasma Parent Drug Concentration - \n Time Profile by Age (260 mg)")
  cat('&nbsp;\n\n') 
t4 <- all3trial |> 
  filter(StudyID==3, DoseAMT==390, Time %in% c(0, 24, 48, 72, 96, 120, 144)) |> 
  ggplot(aes(x=Time, y=C1Obs, colour = factor(ID))) +
  geom_point(size= 3, alpha=0.3)+
  geom_line(size = .5)+
  xlab("Time (Hours)") + ylab("Plasma Concentration (mg/L)") +
  ggtitle("Plasma Parent Drug Concentration - \n Time Profile (390 mg)")

  t4 + facet_wrap(~Sex, ncol=2) +
  ggtitle("Plasma Parent Drug Concentration - \n Time Profile by Sex (390 mg)")
  t4 + facet_wrap(~factor(AgeBins), ncol=2) +
  ggtitle("Plasma Parent Drug Concentration - \n Time Profile by Age (390 mg)")
 
  t1/t2
  
cat('&nbsp;\n\n')
cat('&nbsp;\n\n')
  
  t3/t4
  
cat('&nbsp;\n\n')
cat('\\newpage\n\n')
```
# Non-Linear Regression Modeling
## Study 1
```{r Modeling}

#for Study1
all3trial_study1 <- all3trial |> 
  filter(StudyID == 1)
C1Obs <- all3trial_study1$C1Obs
Time <- all3trial_study1$Time
#find the parameters for the equation
SS<-getInitial(C1Obs~SSlogis(Time,alpha,xmid,scale),
               data=data.frame(C1Obs=C1Obs,Time=Time))

K_start<-SS["alpha"]
R_start<-1/SS["scale"]
N0_start<-SS["alpha"]/(exp(SS["xmid"]/SS["scale"])+1)
#the formula for the model
log_formula<-formula(C1Obs~K*N0*exp(R*Time)/(K+N0*(exp(R*Time)-1)), data=all3trial_study1)
#fit the model
model<-nls(log_formula,start=list(K=K_start,R=R_start,N0=N0_start))
#estimated parameters
summary(model)
#Formula: C1Obs ~ K * N0 * exp(R * Time)/(K + N0 * (exp(R * Time) - 1))

#get some estimation of goodness of fit
cor(C1Obs,predict(model))
#[1] 0.6132112

#plot
C1Obs_Study1<- predict(model)
plot(Time,C1Obs_Study1, type = "b", col="red",lty=2,lwd=3)



```